### 2014-11-03 report/preamble.tex
Move preamble to a separate file (finally)

% set fonts for nicer pdf view
\IfFileExists{lmodern.sty}{\usepackage{lmodern}}{}

\fancyheadoffset{12pt}
\renewcommand{\sectionmark}[1]{} % Don't use sections in headers
\renewcommand{\subsectionmark}[1]{} % Don't use subsections in headers

\usepackage{pdflscape}
\usepackage{geometry}

\newenvironment{wide}{
\pagebreak
\newgeometry{
 top=15mm,
 inner=15mm,
 outer=15mm,
 bottom=15mm,
 headheight=15pt,
 headsep=5mm,
 footnotesep=5mm,
 footskip=10mm}
\begin{landscape}
% Use TiKZ?
% http://tex.stackexchange.com/questions/40501/using-restoregeometry-in-environment-next-page-runs-off-the-page-bottom/40503#40503
\setlength\LTcapwidth{\textwidth} % default: 4in (rather less than \textwidth...)
\setlength\LTleft{0pt}            % default: \parindent
\setlength\LTright{0pt}           % default: \fill
}{
\end{landscape}
\pagebreak
\aftergroup\restoregeometry
}

\usepackage{color}
\definecolor{updated}{rgb}{0.8,0.85,1}

\usepackage{numprint}

\providecommand{\versiononly}{(Unknown~document~version)}
\providecommand{\versionandtoday}{\versiononly~--~\today}

\providecommand{\recentlyupdated}{\emph\textsc\textcolor{updated}{ ~Updated!~ }}

\usepackage{longtable}
\usepackage{caption}
\usepackage{pgfplotstable}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{array}
\pgfplotsset{compat=1.11}% supress warning

\pgfplotstableset{
% Write all longtable tables to disk.
every table/.append style={outfile={.generated.report-table\@arabic\c@LT@tables.tex}},
% Use cached/written files as input.
include outfiles,
% Change to true to force re-generation. It's easier to just delete all generated files.
force remake=false,
}

\newcolumntype{i}{>{\begin{pgfplotstablecoltype}[int detect]}r<{\end{pgfplotstablecoltype}}}
\newcolumntype{f}{>{\begin{pgfplotstablecoltype}[fixed zerofill]}r<{\end{pgfplotstablecoltype}}}

\newcommand{\tsvtablestyle}[0]{
\pgfplotstableset{
format=file,
col sep=tab,
multicolumn names={c},
column type=r,
int detect,
zerofill=true,
precision=3,
1000 sep={\ },
begin table=\begin{longtable*},
end table=\end{longtable*},
every odd row/.style={before row={\rowcolor[gray]{0.95}}},
every head row/.style={before row=\toprule,after row={
\toprule
\addlinespace[0pt]
\endhead
},
},
every last row/.style={after row={
\bottomrule
}
},
}
}

\newcommand{\tsvtable}[3]{
\tsvtablestyle
\pgfplotstabletypeset[
columns={#2},#3
]{#1}
}

\newcommand{\tsvtablestylewide}[1]{
\pgfplotstableset{
format=file,
col sep=tab,
multicolumn names={c},
column type=r,
int detect,
zerofill=true,
precision=3,
1000 sep={\ },
begin table=\begin{longtable},
end table=\end{longtable},
every odd row/.style={before row={\rowcolor[gray]{0.95}}},
every head row/.style={before row=\toprule,after row={
\toprule
\addlinespace[0pt]
\endhead
},
},
every last row/.style={after row={
\bottomrule
\caption{#1}}
},
}
}

\newcommand{\tsvtablewide}[4]{
\tsvtablestylewide{#2}
\pgfplotstabletypeset[
columns={#3},#4
]{#1}
}

\newcommand{\tsvtablewidedatasets}[4]{
\tsvtablestylewide{#2}
\pgfplotstabletypeset[
columns={#3},every nth row={4[-1]}{after row={\addlinespace[0pt]\midrule\addlinespace[0pt]}},fixed,display columns/0/.style={string type, column type=l},display columns/1/.style={string type, column type=i},#4
]{#1}
}

\newenvironment{futurework}{
\begin{center}
\begin{minipage}{.8\columnwidth}
\hspace{12pt}
\hrule
\hspace{12pt}
}{
\hspace{12pt}
\hrule
\hspace{12pt}
\end{minipage}
\end{center}
}

% Speed up pgfplots/tikz figure generation by caching results.
% LyX needs to be configured per user;
% In LyX Preferences, File Handling, Converters:
% For LaTeX (pdflatex) -> PDF (pdflatex) set the Converted field to:
% pdflatex -shell-escape $$i
% (Adding  -shell-escape is what is important.)
\usepgfplotslibrary{external}
% When using LyX, the command is executed in a temporary working directory; can't use a folder prefix.
\tikzexternalize[prefix=.generated.]
% Use to remake all or remake a single figure.
%\tikzset{external/force remake}
%\tikzset{external/remake next}

\pgfplotsset{allDatasetsStackedXbar/.style=
{	
xbar stacked=plus,
enlarge x limits=false,
bar width=6pt,
fill opacity=0.9,
xmin=0,
xmax=1,
xtick={0,.2,.4,.6,.8,1},
minor xtick={.1,.3,.5,.7,.9},
xminorgrids=true,
xmajorgrids=true,
% The number style seems to have stopped working?
xticklabel style={
/pgf/number format/fixed,
/pgf/number format/precision=2
},
ytick=data,
y dir=reverse,
yticklabel pos=right,
symbolic y coords={alexa.2014-09-01.random.10000-http,alexa.2014-09-01.random.10000-http-www,alexa.2014-09-01.random.10000-https,alexa.2014-09-01.random.10000-https-www,alexa.2014-09-01.top.10000-http,alexa.2014-09-01.top.10000-http-www,alexa.2014-09-01.top.10000-https,alexa.2014-09-01.top.10000-https-www,alexa.2014-09-01.top.dk.10000-http,alexa.2014-09-01.top.dk.10000-http-www,alexa.2014-09-01.top.dk.10000-https,alexa.2014-09-01.top.dk.10000-https-www,alexa.2014-09-01.top.se.10000-http,alexa.2014-09-01.top.se.10000-http-www,alexa.2014-09-01.top.se.10000-https,alexa.2014-09-01.top.se.10000-https-www,com.2014-08-29.random.10000-http,com.2014-08-29.random.10000-http-www,com.2014-08-29.random.10000-https,com.2014-08-29.random.10000-https-www,dk.2014-07-23.random.10000-http,dk.2014-07-23.random.10000-http-www,dk.2014-07-23.random.10000-https,dk.2014-07-23.random.10000-https-www,net.2014-08-29.random.10000-http,net.2014-08-29.random.10000-http-www,net.2014-08-29.random.10000-https,net.2014-08-29.random.10000-https-www,reach50.2014w35.se-http,reach50.2014w35.se-http-www,reach50.2014w35.se-https,reach50.2014w35.se-https-www,se.2014-07-10.random.100000-http,se.2014-07-10.random.100000-http-www,se.2014-07-10.random.100000-https,se.2014-07-10.random.100000-https-www,se.healthstatus.2013.counties-http,se.healthstatus.2013.counties-http-www,se.healthstatus.2013.counties-https,se.healthstatus.2013.counties-https-www,se.healthstatus.2013.domain-registrars-http,se.healthstatus.2013.domain-registrars-http-www,se.healthstatus.2013.domain-registrars-https,se.healthstatus.2013.domain-registrars-https-www,se.healthstatus.2013.financial-services-http,se.healthstatus.2013.financial-services-http-www,se.healthstatus.2013.financial-services-https,se.healthstatus.2013.financial-services-https-www,se.healthstatus.2013.gocs-http,se.healthstatus.2013.gocs-http-www,se.healthstatus.2013.gocs-https,se.healthstatus.2013.gocs-https-www,se.healthstatus.2013.higher-education-http,se.healthstatus.2013.higher-education-http-www,se.healthstatus.2013.higher-education-https,se.healthstatus.2013.higher-education-https-www,se.healthstatus.2013.isps-http,se.healthstatus.2013.isps-http-www,se.healthstatus.2013.isps-https,se.healthstatus.2013.isps-https-www,se.healthstatus.2013.media-http,se.healthstatus.2013.media-http-www,se.healthstatus.2013.media-https,se.healthstatus.2013.media-https-www,se.healthstatus.2013.municipalities-http,se.healthstatus.2013.municipalities-http-www,se.healthstatus.2013.municipalities-https,se.healthstatus.2013.municipalities-https-www,se.healthstatus.2013.public-authorities-http,se.healthstatus.2013.public-authorities-http-www,se.healthstatus.2013.public-authorities-https,se.healthstatus.2013.public-authorities-https-www},
legend image code/.code={%
\draw[#1] (0pt,-4pt) rectangle (8pt,4pt);
},
},
allDatasetsStackedXbarFullPage/.style=
{	
enlarge y limits=0.006,
width=.6\textwidth,
height=1.4\textwidth,
font=\small
}
}

\pgfplotsset{datasetsCDF/.style=
{	
xmin=0,
xmax=1,
ymin=0,
ymax=1,
grid=both,
xtick={0,.1,...,1.1},
ytick={0,.1,...,1.1},
enlargelimits=false,
% Marks 0% and 99%.
mark repeat=99,
legend pos=south east,
legend style={
cells={anchor=west},
},
},
datasetsCDFFullPage/.style=
{	
width=\textwidth,
height=1.425\textwidth,
}
}