### 2014-11-17 report/preamble.tex
Add result summary figures

% set fonts for nicer pdf view
\IfFileExists{lmodern.sty}{\usepackage{lmodern}}{}

\fancyheadoffset{12pt}
\renewcommand{\sectionmark}[1]{} % Don't use sections in headers
\renewcommand{\subsectionmark}[1]{} % Don't use subsections in headers

\usepackage{pdflscape}
\usepackage{geometry}

\newenvironment{wide}{
\pagebreak
\newgeometry{
 top=15mm,
 inner=15mm,
 outer=15mm,
 bottom=15mm,
 headheight=15pt,
 headsep=5mm,
 footnotesep=5mm,
 footskip=10mm}
\begin{landscape}
% Use TiKZ?
% http://tex.stackexchange.com/questions/40501/using-restoregeometry-in-environment-next-page-runs-off-the-page-bottom/40503#40503
\setlength\LTleft{0pt}            % default: \fill
\setlength\LTright{0pt}           % default: \fill
}{
\end{landscape}
\pagebreak
\aftergroup\restoregeometry
}

\newenvironment{portraitdatapage}{
\pagebreak
\newgeometry{
 top=15mm,
 inner=15mm,
 outer=15mm,
 bottom=15mm,
 headheight=15pt,
 headsep=5mm,
 footnotesep=5mm,
 footskip=10mm}
% Use TiKZ?
% http://tex.stackexchange.com/questions/40501/using-restoregeometry-in-environment-next-page-runs-off-the-page-bottom/40503#40503
\setlength\LTleft{0pt}            % default: \fill
\setlength\LTright{0pt}           % default: \fill
}{
\pagebreak
\aftergroup\restoregeometry
}

\usepackage{color}
\definecolor{updated}{rgb}{0.8,0.85,1}

\usepackage{url}
\usepackage{numprint}

\newcommand{\readsinglelinefile}[1]{
	\newread\inputstream%
	\immediate\openin\inputstream=#1%
	\immediate\read\inputstream to \singlelinecontents%
	\singlelinecontents%
	\immediate\closein\inputstream%
}

\providecommand{\versiononly}{\IfFileExists{.generated.version}{\readsinglelinefile{.generated.version}}{(Unknown~document~version)}}%
\providecommand{\versionandtoday}{\versiononly~--~\today}

\providecommand{\recentlyupdated}{\emph\textsc\textcolor{updated}{ ~Updated!~ }}

\usepackage{longtable}
\setlength\LTcapwidth{\textwidth} % default: 4in (rather less than \textwidth...)

\usepackage{caption}
\usepackage{pgfplotstable}
\usepackage{booktabs}
\usepackage{colortbl}
\usepackage{array}
\pgfplotsset{compat=1.11}% supress warning

\pgfplotstableset{
% Write all longtable tables to disk.
every table/.append style={outfile={.generated.report-table\@arabic\c@LT@tables.tex}},
% Use cached/written files as input.
include outfiles,
% Change to true to force re-generation. It's easier to just delete all generated files.
% force remake=true,
}

\newcolumntype{i}{>{\begin{pgfplotstablecoltype}[int detect]}r<{\end{pgfplotstablecoltype}}}
\newcolumntype{f}{>{\begin{pgfplotstablecoltype}[fixed zerofill]}r<{\end{pgfplotstablecoltype}}}
\newcolumntype{t}{>{\arraybackslash}p{.3\linewidth}}
\newcolumntype{T}{>{\arraybackslash}p{.9\linewidth}}

% TODO: reduce code-duplication between \tsvtablestyle/wide/widedatasets/superwidedatasets
\newcommand{\tsvtablestyle}[0]{
\pgfplotstableset{
format=file,
col sep=tab,
multicolumn names={c},
column type=r,
int detect,
zerofill=true,
precision=3,
1000 sep={\ },
begin table=\begin{longtable*},
end table=\end{longtable*},
every odd row/.style={before row={\rowcolor[gray]{0.95}}},
every head row/.style={before row=\toprule,after row={
\toprule
\addlinespace[0pt]
\endhead
},
},
every last row/.style={after row={
\bottomrule
}
},
}
}

\newcommand{\tsvtable}[3]{
\tsvtablestyle
\pgfplotstabletypeset[
columns={#2},#3
]{#1}
}

\newcommand{\tsvtablestylewide}[1]{
\pgfplotstableset{
format=file,
col sep=tab,
multicolumn names={c},
column type=r,
int detect,
zerofill=true,
precision=3,
1000 sep={\ },
begin table=\begin{longtable},
end table=\end{longtable},
every odd row/.style={before row={\rowcolor[gray]{0.95}}},
every head row/.style={before row=\toprule,after row={
\toprule
\addlinespace[0pt]
\endhead
},
},
every last row/.style={after row={
\bottomrule
\caption{#1}}
},
}
}

\newcommand{\tsvtablewide}[4]{
\tsvtablestylewide{#2}
\pgfplotstabletypeset[
columns={#3},#4
]{#1}
}

\newcommand{\tsvtablewidedatasets}[4]{
\tsvtablestylewide{#2}
\pgfplotstabletypeset[
columns={#3},every nth row={4[-1]}{after row={\addlinespace[0pt]\midrule\addlinespace[0pt]}},fixed,display columns/0/.style={string type, column type=l},display columns/1/.style={string type, column type=i|},#4
]{#1}
}

\newcommand{\tsvtablesuperwidedatasets}[4]{
\tsvtablestylewide{#2}
\pgfplotstabletypeset[
columns={#3},every nth row={4[-1]}{after row={\addlinespace[0pt]\midrule\addlinespace[0pt]}},fixed,display columns/0/.style={string type, column type=l},display columns/1/.style={string type, column type=i|},%
% Remove dataset dates.
display columns/0/.append style={string replace*={2014-09-01.}{}},%
display columns/0/.append style={string replace*={2014-08-29.}{}},%
display columns/0/.append style={string replace*={2014-07-23.}{}},%
display columns/0/.append style={string replace*={2014w35.}{}},%
display columns/0/.append style={string replace*={2014-07-10.}{}},%
% Shorten health status datasets.
display columns/0/.append style={string replace*={healthstatus.2013.}{hs.}},%
display columns/0/.append style={string replace*={.domain-registrars}{.registrars}},%
display columns/0/.append style={string replace*={.financial-services}{.financial}},%
display columns/0/.append style={string replace*={.higher-education}{.education}},%
display columns/0/.append style={string replace*={.public-authorities}{.pubauth}},%
% Shorten dataset counts.
display columns/0/.append style={string replace*={100000}{100k}},%
display columns/0/.append style={string replace*={10000}{10k}},%
% Shorten Alexa datasets.
display columns/0/.append style={string replace*={alexa.top.se.10k.}{alexa.top.se.}},%
display columns/0/.append style={string replace*={alexa.top.dk.10k.}{alexa.top.dk.}},%
% Shorten dataset type.
display columns/0/.append style={string replace*={random.}{rnd.}},%
display columns/0/.append style={string replace*={-https-www}{-sw}},%
display columns/0/.append style={string replace*={-https}{-s}},%
display columns/0/.append style={string replace*={-http-www}{-hw}},%
display columns/0/.append style={string replace*={-http}{-h}},%
,#4
]{#1}
}

\newenvironment{futurework}{
\begin{center}
\begin{minipage}{.8\columnwidth}
\hspace{12pt}
\hrule
\hspace{12pt}
}{
\hspace{12pt}
\hrule
\hspace{12pt}
\end{minipage}
\end{center}
}

% Speed up pgfplots/tikz figure generation by caching results.
% LyX needs to be configured per user;
% In LyX Preferences, File Handling, Converters:
% For LaTeX (pdflatex) -> PDF (pdflatex) set the Converted field to:
% pdflatex -shell-escape $$i
% (Adding  -shell-escape is what is important.)
\usepgfplotslibrary{external}
% When using LyX, the command is executed in a temporary working directory; can't use a folder prefix.
\tikzexternalize[prefix=.generated.]
% Use to remake all or remake a single figure.
%\tikzset{external/force remake}
%\tikzset{external/remake next}

\usepgfplotslibrary{groupplots}

\pgfplotsset{allDatasetsXbar/.style=
{
enlarge x limits=false,
bar width=6pt,
fill opacity=0.9,
xmin=0,
xmax=1,
xtick={0,.2,.4,.6,.8,1},
minor xtick={.1,.3,.5,.7,.9},
xminorgrids=true,
xmajorgrids=true,
ytick=data,
y dir=reverse,
yticklabel pos=right,
symbolic y coords={alexa.2014-09-01.random.10000-http,alexa.2014-09-01.random.10000-http-www,alexa.2014-09-01.random.10000-https,alexa.2014-09-01.random.10000-https-www,alexa.2014-09-01.top.10000-http,alexa.2014-09-01.top.10000-http-www,alexa.2014-09-01.top.10000-https,alexa.2014-09-01.top.10000-https-www,alexa.2014-09-01.top.dk.10000-http,alexa.2014-09-01.top.dk.10000-http-www,alexa.2014-09-01.top.dk.10000-https,alexa.2014-09-01.top.dk.10000-https-www,alexa.2014-09-01.top.se.10000-http,alexa.2014-09-01.top.se.10000-http-www,alexa.2014-09-01.top.se.10000-https,alexa.2014-09-01.top.se.10000-https-www,com.2014-08-29.random.10000-http,com.2014-08-29.random.10000-http-www,com.2014-08-29.random.10000-https,com.2014-08-29.random.10000-https-www,dk.2014-07-23.random.10000-http,dk.2014-07-23.random.10000-http-www,dk.2014-07-23.random.10000-https,dk.2014-07-23.random.10000-https-www,net.2014-08-29.random.10000-http,net.2014-08-29.random.10000-http-www,net.2014-08-29.random.10000-https,net.2014-08-29.random.10000-https-www,reach50.2014w35.se-http,reach50.2014w35.se-http-www,reach50.2014w35.se-https,reach50.2014w35.se-https-www,se.2014-07-10.random.100000-http,se.2014-07-10.random.100000-http-www,se.2014-07-10.random.100000-https,se.2014-07-10.random.100000-https-www,se.healthstatus.2013.counties-http,se.healthstatus.2013.counties-http-www,se.healthstatus.2013.counties-https,se.healthstatus.2013.counties-https-www,se.healthstatus.2013.domain-registrars-http,se.healthstatus.2013.domain-registrars-http-www,se.healthstatus.2013.domain-registrars-https,se.healthstatus.2013.domain-registrars-https-www,se.healthstatus.2013.financial-services-http,se.healthstatus.2013.financial-services-http-www,se.healthstatus.2013.financial-services-https,se.healthstatus.2013.financial-services-https-www,se.healthstatus.2013.gocs-http,se.healthstatus.2013.gocs-http-www,se.healthstatus.2013.gocs-https,se.healthstatus.2013.gocs-https-www,se.healthstatus.2013.higher-education-http,se.healthstatus.2013.higher-education-http-www,se.healthstatus.2013.higher-education-https,se.healthstatus.2013.higher-education-https-www,se.healthstatus.2013.isps-http,se.healthstatus.2013.isps-http-www,se.healthstatus.2013.isps-https,se.healthstatus.2013.isps-https-www,se.healthstatus.2013.media-http,se.healthstatus.2013.media-http-www,se.healthstatus.2013.media-https,se.healthstatus.2013.media-https-www,se.healthstatus.2013.municipalities-http,se.healthstatus.2013.municipalities-http-www,se.healthstatus.2013.municipalities-https,se.healthstatus.2013.municipalities-https-www,se.healthstatus.2013.public-authorities-http,se.healthstatus.2013.public-authorities-http-www,se.healthstatus.2013.public-authorities-https,se.healthstatus.2013.public-authorities-https-www},
legend image code/.code={%
\draw[#1] (0pt,-4pt) rectangle (8pt,4pt);
},
},
allDatasetsXbarFullPage/.style=
{
enlarge y limits=0.008,
width=.65\textwidth,
height=1.47\textwidth,
font=\small,
legend columns=-1,
legend style={at={(0.5,-16pt)},anchor=north}
},
allDatasetsXbarMiniResults/.style=
{
enlarge y limits=0.1,
width=.35\textwidth,
height=.3\textwidth,
font=\tiny,
legend columns=-1,
legend style={at={(0.5,-12pt)},anchor=north},
symbolic y coords={alexa.2014-09-01.top.10000-http-www,alexa.2014-09-01.top.10000-https-www,alexa.2014-09-01.top.se.10000-http-www,alexa.2014-09-01.top.se.10000-https-www,se.2014-07-10.random.100000-http-www,se.2014-07-10.random.100000-https-www,se.healthstatus.2013.municipalities-http-www,se.healthstatus.2013.municipalities-https-www,},
xticklabel style={
/pgf/number format/fixed,
/pgf/number format/fixed zerofill,
/pgf/number format/precision=1
},
title style={at={(.5,.95)}},
},
xticklabel style={
/pgf/number format/fixed,
/pgf/number format/fixed zerofill,
/pgf/number format/precision=2
},
yticklabel style={
/pgf/number format/fixed,
/pgf/number format/fixed zerofill,
/pgf/number format/precision=2
}
}
\pgfplotsset{allDatasetsStackedXbar/.style=
{
xbar stacked=plus,
},
}

\pgfplotsset{allDatasetsMultiXbar/.style=
{
xbar,
},
allDatasetsMultiXbarPart2/.style=
{
bar width=4pt,
% bar direction=y,
very thin,
ymajorgrids=true,
major y grid style={loosely dashed},
},
}

\newcommand{\XBarPlotFromDataset}[5]{
	\addplot[
			fill=#4,#5
		]
		table[y=Dataset,x=#2]
		{#1};
	\addlegendentry{#3}
}

\pgfplotsset{datasetsCDF/.style=
{
xmin=0,
xmax=1,
ymin=0,
ymax=1,
grid=both,
xtick={0,.1,...,1.1},
ytick={0,.1,...,1.1},
enlargelimits=false,
% Marks 0% and 99%.
mark repeat=99,
% Using a custom position below.
% legend pos=south east,
legend style={
at={(.99,.01)},
anchor=south east,
cells={anchor=west},
},
},
datasetsCDFFullPage/.style=
{
width=\textwidth,
height=1.46\textwidth,
}
}

\pgfplotsset{datasetsOccurances/.style=
{
ymin=,
xmin=0.1,
xmax=100,
xtick={0.1,1,2,3,4,5,10,20,50,100},
minor xtick={6,7,8,9,30,40,60,70,80,90},
xticklabel={%
	\ifnum\ticknum=9%
		$100^+$%
	\else%
		\pgfmathfloatparsenumber{\tick}%
		\pgfmathfloatexp{\pgfmathresult}%
		\pgfmathprintnumber[fixed,precision=0]{\pgfmathresult}%
	\fi%
},
yticklabel style={
/pgf/number format/fixed,
/pgf/number format/fixed zerofill,
/pgf/number format/precision=2
},
% Marks 0.1.
mark repeat=1,
mark indices={1},
}
}

\newcommand{\CDFPlotFromDataset}[3]{
	\addplot+[#3,solid,mark=*,every mark/.append style={#3,solid},]
		table[x=colnames,y=#2-http-www]
		{#1};
	\addlegendentry{#2-http-www}
	\addplot+[#3,dashed,mark=square*,every mark/.append style={#3,solid},]
		table[x=colnames,y=#2-https-www]
		{#1};
	\addlegendentry{#2-https-www}
}

\newcommand{\CDFPlotFromDefaultDatasets}[1]{
	\CDFPlotFromDataset{#1}{alexa.2014-09-01.top.10000}{red}
	\CDFPlotFromDataset{#1}{alexa.2014-09-01.top.se.10000}{blue}
	\CDFPlotFromDataset{#1}{se.2014-07-10.random.100000}{brown}
	\CDFPlotFromDataset{#1}{se.healthstatus.2013.municipalities}{violet}
	% \CDFPlotFromDataset{#1}{se.healthstatus.2013.media}{teal}
}

\newcommand{\CDFPlotFromDefaultDatasetsFile}[1]{
	\pgfplotstabletranspose[colnames from=Dataset]\tableTransposedDatasetsFromFile{#1}

	\CDFPlotFromDefaultDatasets{\tableTransposedDatasetsFromFile}
}