### 2015-02-03 report/tableread.tex
Write about undetected external domains, add table and graph. Add "the end" page.

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{disconnect.domains-per-organization.tsv}\tableDisconnectDomainsPerOrganization

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.request-status.coverage.origin.sorted.tsv}\tableDatasetsRequestStatusCoverageOriginSorted

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.classification.domain-scope.coverage.sorted.tsv}\tableDatasetsNonFailedClassificationDomainScopeCoverageSorted

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.classification.secure.coverage.sorted.tsv}\tableDatasetsNonFailedClassificationSecureCoverageSorted

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.origin-redirects.coverage.sorted.tsv}\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.disconnect.categories.coverage.external.sorted.tsv}\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.disconnect.organizations.coverage.external.sorted.tsv}\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSorted

\pgfplotstabletranspose[
	colnames from=Dataset
	]\tableTransposedDatasetsNonFailedRatioBucketsIsDisconnectMatchNormalizedCumulativeSorted{datasets.non-failed.ratio-buckets.is-disconnect-match.normalized.cumulative.sorted.tsv}

\pgfplotstabletranspose[
	colnames from=Dataset
	]\tableTransposedDatasetsNonFailedRatioBucketsDisconnectOrganizationsNormalizedCumulativeSorted{datasets.non-failed.ratio-buckets.disconnect-organizations.normalized.cumulative.sorted.tsv}

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.domains.ratios.sorted.tsv}\tableDatasetsNonFailedDomainsRatiosSorted

% Summaries
% TODO: see if it's possible to filter rows from tables loaded above
\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.request-status.coverage.origin.sorted.summary.tsv}\tableDatasetsRequestStatusCoverageOriginSortedSummary

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.classification.domain-scope.coverage.sorted.summary.tsv}\tableDatasetsNonFailedClassificationDomainScopeCoverageSortedSummary

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.origin-redirects.coverage.sorted.summary.tsv}\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSummary

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.classification.secure.coverage.sorted.summary.tsv}\tableDatasetsNonFailedClassificationSecureCoverageSortedSummary

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.disconnect.organizations.coverage.external.sorted.summary.tsv}\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSummary

\pgfplotstableread[
		format=file,
		col sep=tab,
	]{datasets.non-failed.disconnect.categories.coverage.external.sorted.summary.tsv}\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary