### 2015-04-28 paper/preamble.tex
Move page and table commands out of the preamble to allow externalized generation

% set fonts for nicer pdf view
\IfFileExists{lmodern.sty}{\usepackage{lmodern}}{}

\fancyheadoffset{12pt}
\renewcommand{\sectionmark}[1]{} % Don't use sections in headers
\renewcommand{\subsectionmark}[1]{} % Don't use subsections in headers

\usepackage{url}
\usepackage{numprint}
\npstyleenglish

\newcommand{\readsinglelinefile}[1]{
	\newread\inputstream%
	\immediate\openin\inputstream=#1%
	\immediate\read\inputstream to \singlelinecontents%
	\singlelinecontents%
	\immediate\closein\inputstream%
}

\providecommand{\versiononly}{\IfFileExists{.generated.version}{\readsinglelinefile{.generated.version}}{(Unknown~document~version)}}%
\providecommand{\versionandtoday}{\versiononly~--~\today}

\newenvironment{futurework}{
\begin{center}
\begin{minipage}{.8\columnwidth}
\hspace{12pt}
\hrule
\hspace{12pt}
}{
\hspace{12pt}
\hrule
\hspace{12pt}
\end{minipage}
\end{center}
}

\usepackage{jppagecommands}
\usepackage[pfx=.generated.paper-]{jptablecommands}
\usepackage[pfx=.generated.]{jpfigurecommands}