### 2016-08-14 paper/figures.tex
Render some bar graphs in grayscale

\documentclass[a4paper,english]{report}

\include{rendering-preamble}

\usepackage[pfx=figures-]{jpfigurecommands}
\include{tableread}

\begin{document}

\tikzsetnextfilename{medium-domains-per-organization}
\begin{tikzpicture}
\begin{loglogaxis}[
	xlabel=Domains,
	ylabel=Organizations,
	width=\textwidth,
	height=0.5*\textwidth,
	log ticks with fixed point,
	xmin=0.7,
	ymin=0.7,
	enlarge x limits=upper,
	enlarge y limits=upper,
	xtick={1,2,3,4,10,32,40,71,271},
	ytick=data,
]
\addplot[
		ybar,
		bar width=3pt,
		thick,
		fill=black,
	]
	table[x=Domains per organization,y=Organizations]
	\tableDisconnectDomainsPerOrganization;
\end{loglogaxis}
\end{tikzpicture}

\tikzsetnextfilename{mini-3x1-status-redirects-secure-xbar}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=3 by 1,
		horizontal sep=8pt,
	},
]

\nextgroupplot[
	% title=(a) Origin HTTP response codes,
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	yticklabels={},
]

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSortedSummary}{2xx}{2xx}{green}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSortedSummary}{3xx}{3xx}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSortedSummary}{(null)}{(null)}{red}{}

\nextgroupplot[
	% title=(b) Origin redirects,
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	yticklabels={},
	legend columns=2,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSummary}{Sec}{Secure}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSummary}{Mix sec}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSummary}{Insec}{Insecure}{orange}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSummary}{Mism}{Mismatch}{lightgray}{}

\nextgroupplot[
	% title=(c) Resource security,
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	yticklabels={Top 10k global,Top 10k global (HTTPS), Top 10k .se,Top 10k .se (HTTPS),Rand 100k .se,Rand 100k .se (HTTPS),Municipal .se,Municipal .se (HTTPS)},
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSortedSummary}{All sec}{Secure}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSortedSummary}{Mix sec}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSortedSummary}{All insec}{Insecure}{orange}{}

\end{groupplot}
\end{tikzpicture}

\tikzsetnextfilename{mini-3x1-internal-categories-organizations-xbar}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=3 by 1,
		horizontal sep=8pt,
	},
]

\nextgroupplot[
	title=(a) Resource distribution,
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	yticklabels={},
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSortedSummary}{Internal}{Internal}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSortedSummary}{Mixed}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSortedSummary}{External}{External}{orange}{}

\nextgroupplot[
	title=(b) Tracker categories,
	allDatasetsMultiXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	allDatasetsMultiXbarPart2,
	bar width=2.5pt,
	yticklabels={},
	legend columns=3,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\nextgroupplot[
	title=(c) Top organizations,
	allDatasetsMultiXbar,
	allDatasetsXbar,
	allDatasetsXbarMiniResults,
	allDatasetsMultiXbarPart2,
	bar width=2.5pt,
	yticklabels={alexa.top.10k-hw,alexa.top.10k-sw,alexa.top.se-hw,alexa.top.se-sw,se.r.100k-hw,se.r.100k-sw,se.hs.municip-hw,se.hs.municip-sw},
	legend columns=3,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary}{Any}{Any}{black!20!white}{bar shift=0pt, bar width=8.5pt}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSummary}{Google}{Google}{red!90!white}{ultra thin, bar shift=2pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSummary}{Facebook}{Facebook}{blue!90!white}{ultra thin, bar shift=0pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSummary}{Twitter}{Twitter}{cyan!70!white}{ultra thin, bar shift=-2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Dataset,x=Disconnect]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSummary;
\addlegendentry{Disconnect}

\end{groupplot}
\end{tikzpicture}

\tikzsetnextfilename{mini-3x1-internal-secure-organizations-cdf}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=3 by 1,
		horizontal sep=8pt,
	},
]

\nextgroupplot[
	title=(a) Internal resources,
	datasetsCDF,
	datasetsCDFMiniResults,
	xlabel=Ratio of internal resources,
	ylabel=Ratio of domains,
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-internal-domain.normalized.cumulative.sorted.tsv}

\legend{}

\nextgroupplot[
	title=(b) Secure resources,
	datasetsCDF,
	datasetsCDFMiniResults,
	xlabel=Ratio of secure resources,
	%ylabel=Ratio of domains,
	yticklabels={},
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-secure.normalized.cumulative.sorted.tsv}

\legend{}

\nextgroupplot[
	title=(c) Disconnect's organizations,
	datasetsCDF,
	datasetsOccurances,
	datasetsOccurancesMiniResults,
	xlabel=Number of organizations,
	%ylabel=Ratio of domains,
	yticklabels={},
	legend columns=1,
	legend style={at={(1.025,.5)}, anchor=west},
]

	\CDFPlotFromDefaultDatasets{\tableTransposedDatasetsNonFailedRatioBucketsDisconnectOrganizationsNormalizedCumulativeSorted}

\legend{alexa.top.10k-hw,alexa.top.10k-sw,alexa.top.se-hw,alexa.top.se-sw,se.r.100k-hw,se.r.100k-sw,se.hs.municip-hw,se.hs.municip-sw}

\end{groupplot}
\end{tikzpicture}

\tikzsetnextfilename{full-http-status-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
]

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{1xx}{1xx}{orange}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{2xx}{2xx}{green}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{3xx}{3xx}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{4xx}{4xx}{orange}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{5xx}{5xx}{orange}{}

\XBarPlotFromDataset{\tableDatasetsRequestStatusCoverageOriginSorted}{(null)}{(null)}{red}{}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-internal-resources-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSorted}{Internal}{Internal}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSorted}{Mixed}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationDomainScopeCoverageSorted}{External}{External}{orange}{}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-internal-resources-cdf}
\begin{tikzpicture}
\begin{axis}[
	datasetsCDF,
	datasetsCDFFullPage,
	xlabel=Ratio of internal resources,
	ylabel=Ratio of domains,
	legend style={
		at={(.01,.99)},
		anchor=north west,
	},
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-internal-domain.normalized.cumulative.sorted.tsv}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-secure-resources-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSorted}{All sec}{Secure}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSorted}{Mix sec}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationSecureCoverageSorted}{All insec}{Insecure}{orange}{}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-secure-resources-cdf}
\begin{tikzpicture}
\begin{axis}[
	datasetsCDF,
	datasetsCDFFullPage,
	xlabel=Ratio of secure resources,
	ylabel=Ratio of domains,
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-secure.normalized.cumulative.sorted.tsv}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-origin-redirect-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted}{Sec}{Secure}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted}{Mix sec}{Mixed}{yellow}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted}{Insec}{Insecure}{orange}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted}{Mism}{Mismatch}{lightgray}{}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		thick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Dataset,x=Final sec]
	\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSorted;
\addlegendentry{Final redirect secure}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-disconnect-matches-cdf}
\begin{tikzpicture}
\begin{axis}[
	datasetsCDF,
	datasetsCDFFullPage,
	xlabel=Ratio of Disconnect matches,
	ylabel=Ratio of domains,
]

	\CDFPlotFromDefaultDatasets{\tableTransposedDatasetsNonFailedRatioBucketsIsDisconnectMatchNormalizedCumulativeSorted}

	\CDFPlotFromDataset{\tableTransposedDatasetsNonFailedRatioBucketsIsDisconnectMatchNormalizedCumulativeSorted}{se.healthstatus.2013.media}{teal}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-disconnect-organizations-cdf}
\begin{tikzpicture}
\begin{semilogxaxis}[
	datasetsCDF,
	datasetsOccurances,
	datasetsCDFFullPage,
	xlabel=Number of organizations,
	ylabel=Ratio of domains,
]

	\CDFPlotFromDefaultDatasets{\tableTransposedDatasetsNonFailedRatioBucketsDisconnectOrganizationsNormalizedCumulativeSorted}

	\CDFPlotFromDataset{\tableTransposedDatasetsNonFailedRatioBucketsDisconnectOrganizationsNormalizedCumulativeSorted}{se.healthstatus.2013.media}{teal}

\end{semilogxaxis}
\end{tikzpicture}

\tikzsetnextfilename{full-disconnect-categories-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsMultiXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
	allDatasetsMultiXbarPart2,
	bar width=2.5pt,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-disconnect-top-organizations-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsMultiXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
	allDatasetsMultiXbarPart2,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted}{Any}{Any}{black!20!white}{bar shift=0pt, bar width=8.5pt}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSorted}{Google}{Google}{red!90!white}{ultra thin, bar shift=2pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSorted}{Facebook}{Facebook}{blue!90!white}{ultra thin, bar shift=0pt,}

\XBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSorted}{Twitter}{Twitter}{cyan!70!white}{ultra thin, bar shift=-2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Dataset,x=Disconnect]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSorted;
\addlegendentry{Disconnect}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{full-disconnect-detection-xbar}
\begin{tikzpicture}
\begin{axis}[
	allDatasetsStackedXbar,
	allDatasetsXbar,
	allDatasetsXbarFullPage,
]

\XBarPlotFromDataset{\tableDatasetsNonFailedDomainsRatiosSorted}{Prim. det.}{Detected}{green}{}

\XBarPlotFromDataset{\tableDatasetsNonFailedDomainsRatiosSorted}{Undet.}{Undetected}{orange}{}

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{mini-2x2-categories-organizations-http-ybar}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=2 by 4,
		horizontal sep=4pt,
		vertical sep=8pt,
	},
]

\nextgroupplot[
	% title=B1(a) Tracker categories (http-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	swedenHttpWwwDatasetsYbar,
	width=.60\textwidth,
	xticklabels={},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\legend{}

\nextgroupplot[
	% title=B2(a) Tracker categories (http-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	globalHttpWwwDatasetsYbar,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\nextgroupplot[
	% title=B1(b) Top organizations (http-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	swedenHttpWwwDatasetsYbar,
	width=.60\textwidth,
	xticklabels={Counties,Domain registrars,Financial services,GOCS,Higher education,ISPs,Media,Municipalities,Public authorities,reach50,Top 10k .se,Rand 100k .se,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=5pt,
	},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpWww}{Google}{Google}{black!33!white}{ultra thin, bar shift=-2pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpWww}{Facebook}{Facebook}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpWww}{Twitter}{Twitter}{black!66!white}{ultra thin, bar shift=2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Disconnect,x=Dataset]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpWww;
\addlegendentry{Disconnect}

\legend{}

\nextgroupplot[
	% title=B2(b) Top organizations (http-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	globalHttpWwwDatasetsYbar,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={Top 10k .se,Top 10k .dk,Top 10k global,Rand 100k .se,Rand 10k .dk,Rand 10k .com,Rand 10k .net,Rand 10k global,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=5pt,
	},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpWww}{Google}{Google}{black!33!white}{ultra thin, bar shift=-2pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpWww}{Facebook}{Facebook}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpWww}{Twitter}{Twitter}{black!66!white}{ultra thin, bar shift=2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Disconnect,x=Dataset]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpWww;
\addlegendentry{Disconnect}

\end{groupplot}
\end{tikzpicture}

\tikzsetnextfilename{mini-2x2-categories-organizations-https-ybar}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=2 by 4,
		horizontal sep=4pt,
		vertical sep=8pt,
	},
]

\nextgroupplot[
	% title=B1(a) Tracker categories (https-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	swedenHttpsWwwDatasetsYbar,
	width=.60\textwidth,
	xticklabels={},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\legend{}

\nextgroupplot[
	% title=B2(a) Tracker categories (https-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	globalHttpsWwwDatasetsYbar,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Disconnect}{Disconnect}{black!33!white}{ultra thin, bar shift=-3pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Content}{Content}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=-1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Advertising}{Advertising}{black!66!white}{ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Analytics}{Analytics}{black!10!white}{postaction={pattern=north west lines}, ultra thin, bar shift=1.5pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Social}{Social}{black}{ultra thin, bar shift=3pt,}

\nextgroupplot[
	% title=B1(b) Top organizations (https-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	swedenHttpsWwwDatasetsYbar,
	width=.60\textwidth,
	xticklabels={Counties,Domain registrars,Financial services,GOCS,Higher education,ISPs,Media,Municipalities,Public authorities,reach50,Top 10k .se,Rand 100k .se,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=5pt,
	},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpsWww}{Google}{Google}{black!33!white}{ultra thin, bar shift=-2pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpsWww}{Facebook}{Facebook}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedSwedenHttpsWww}{Twitter}{Twitter}{black!66!white}{ultra thin, bar shift=2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Disconnect,x=Dataset]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedSwedenHttpsWww;
\addlegendentry{Disconnect}

\legend{}

\nextgroupplot[
	% title=B2(b) Top organizations (https-www),
	allDatasetsMultiYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsMultiYbarPart2,
	bar width=2.5pt,
	globalHttpsWwwDatasetsYbar,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={Top 10k .se,Top 10k .dk,Top 10k global,Rand 100k .se,Rand 10k .dk,Rand 10k .com,Rand 10k .net,Rand 10k global,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=5pt,
	},
]

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww}{Any}{Any}{black!02!white}{bar shift=0pt, bar width=8.5pt}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpsWww}{Google}{Google}{black!33!white}{ultra thin, bar shift=-2pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpsWww}{Facebook}{Facebook}{black!10!white}{postaction={pattern=north east lines}, ultra thin, bar shift=0pt,}

\YBarPlotFromDataset{\tableDatasetsNonFailedDisconnectOrganizationsCoverageExternalSortedGlobalHttpsWww}{Twitter}{Twitter}{black!66!white}{ultra thin, bar shift=2pt,}

\addplot+[
		stack plots=false,
		only marks,
		mark=x,
		mark size=3pt,
		fill=white,
		draw=black,
		semithick,
		fill opacity=1,
		update limits=false,
		line legend,
	]
	table[y=Disconnect,x=Dataset]
	\tableDatasetsNonFailedDisconnectCategoriesCoverageExternalSortedGlobalHttpsWww;
\addlegendentry{Disconnect}

\end{groupplot}
\end{tikzpicture}

\createAdjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedHttpsWww{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{\tableDatasetsRequestStatusCoverageOriginSortedSwedenHttpsWww}{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSwedenHttpsWww}
\createAdjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedHttpsWww{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{\tableDatasetsRequestStatusCoverageOriginSortedGlobalHttpsWww}{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedGlobalHttpsWww}
\createAdjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedHttpWww{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpWww}{\tableDatasetsRequestStatusCoverageOriginSortedSwedenHttpWww}{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedSwedenHttpWww}
\createAdjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedHttpWww{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpWww}{\tableDatasetsRequestStatusCoverageOriginSortedGlobalHttpWww}{\tableDatasetsNonFailedClassificationOriginRedirectsCoverageSortedGlobalHttpWww}

\tikzsetnextfilename{mini-2x2-secure-response-redirect-ybar}
\begin{tikzpicture}
\begin{groupplot}[
	group style={
		group size=2 by 2,
		horizontal sep=8pt,
		vertical sep=8pt,
	},
]
\nextgroupplot[
	% title=B1(c) Resp./redir. sec. (https-www),
	allDatasetsStackedYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsStackedYbarPart2,
	swedenHttpsWwwDatasetsYbar,
	width=.60\textwidth,
	xticklabels={},
]

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{2xx}{2xx}{green}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{3xx Sec}{3xx Sec}{green!30!yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{3xx Mix sec}{3xx Mix}{orange}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{3xx Insec}{3xx Insec}{red}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpsWww}{3xx Mism}{3xx Mism}{lightgray}{}

\legend{}

\nextgroupplot[
	% title=B2(c) Resp./redir. sec. (https-www),
	allDatasetsStackedYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsStackedYbarPart2,
	globalHttpsWwwDatasetsYbar,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={},
]

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{2xx}{2xx}{green}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{3xx Sec}{3xx Sec}{green!30!yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{3xx Mix sec}{3xx Mix}{orange}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{3xx Insec}{3xx Insec}{red}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpsWww}{3xx Mism}{3xx Mism}{lightgray}{}

\nextgroupplot[
	% title=B1(d) Resp./redir. sec. (http-www),
	allDatasetsStackedYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsStackedYbarPart2,
	swedenHttpWwwDatasetsYbar,
	paperTight,
	width=.60\textwidth,
	xticklabels={Counties,Domain registrars,Financial services,GOCS,Higher education,ISPs,Media,Municipalities,Public authorities,reach50,Top 10k .se,Rand 100k .se,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=3pt,
	},
]

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpWww}{3xx Sec}{3xx Sec}{green!30!yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpWww}{3xx Mix sec}{3xx Mix}{yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpWww}{3xx Insec}{3xx Insec}{orange}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedSwedenHttpWww}{3xx Mism}{3xx Mism}{lightgray}{}

\legend{}

\nextgroupplot[
	% title=B2(d) Resp./redir. sec. (http-www),
	allDatasetsStackedYbar,
	allDatasetsYbar,
	allDatasetsYbarMiniResults,
	allDatasetsStackedYbarPart2,
	globalHttpWwwDatasetsYbar,
	paperTight,
	width=.44\textwidth,
	yticklabels={},
	xticklabels={Top 10k .se,Top 10k .dk,Top 10k global,Rand 100k .se,Rand 10k .dk,Rand 10k .com,Rand 10k .net,Rand 10k global,},
	xticklabel style={
		rotate=45,
		anchor={north east},
		yshift=3pt,
	},
]

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpWww}{3xx Sec}{3xx Sec}{green!30!yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpWww}{3xx Mix sec}{3xx Mix}{yellow}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpWww}{3xx Insec}{3xx Insec}{orange}{}

\YBarPlotFromDataset{\adjustedTableDatasetsRequestStatusCoverageOriginRedirectSecuritySortedGlobalHttpWww}{3xx Mism}{3xx Mism}{lightgray}{}

\end{groupplot}
\end{tikzpicture}

\tikzsetnextfilename{mini-1x1-internal-resources-cdf}
\begin{tikzpicture}
\begin{axis}[
	% title=(a) Internal resources,
	datasetsCDF,
	datasetsCDFMiniResults,
	paperTight,
	xlabel=Ratio of internal resources,
	ylabel=Ratio of domains,
	legend columns=2,
	legend style={at={(0.5,1.04)}, anchor=south},
	xlabel style={
		at={(0.5,0.1)}
	},
	ylabel style={
		at={(0.17,0.5)}
	},
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-internal-domain.normalized.cumulative.sorted.tsv}

	\legend{Top 10k global,HTTPS, Top 10k .se,HTTPS,Rand 100k .se,HTTPS,Municipal .se,HTTPS},

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{mini-1x1-secure-resources-cdf}
\begin{tikzpicture}
\begin{axis}[
	% title=(b) Secure resources,
	datasetsCDF,
	datasetsCDFMiniResults,
	paperTight,
	xlabel=Ratio of secure resources,
	ylabel=Ratio of domains,
	legend style={at={(1.03,.5)}, anchor=west},
	xlabel style={
		at={(0.5,0.1)}
	},
	ylabel style={
		at={(0.17,0.5)}
	},
]

	\CDFPlotFromDefaultDatasetsFile{datasets.non-failed.ratio-buckets.is-secure.normalized.cumulative.sorted.tsv}

	\legend{Top 10k global,Top 10k global (HTTPS), Top 10k .se,Top 10k .se (HTTPS),Rand 100k .se,Rand 100k .se (HTTPS),Municipal .se,Municipal .se (HTTPS)},

\end{axis}
\end{tikzpicture}

\tikzsetnextfilename{mini-1x1-alexa-rank-plot}
\begin{tikzpicture}
\begin{semilogxaxis}[
	xlabel=Alexa rank,
	ylabel=Organizations,
	xmin=0,
	xmax=1000000,
	ymin=0,
	xmajorgrids=true,
	ymajorgrids=true,
	% xtick={0,10000,100000,1000000},
	xminorgrids=true,
	enlarge x limits=false,
	enlarge y limits=false,
	height=.25\textwidth,
	width=\textwidth,
	legend columns=1,
	legend style={at={(1.05,0.5)},anchor=west,cells={anchor=west},},
	xticklabel style={
		/pgf/number format/.cd,
		/pgf/number format/fixed,
		/pgf/number format/fixed zerofill=false,
		/pgf/number format/precision=2
	},
	yticklabel style={
		/pgf/number format/fixed,
		/pgf/number format/fixed zerofill=false,
		/pgf/number format/precision=2
	},
]
	% \addplot[
	% 	% only marks,
	% 	blue,
	% 	% opacity=0.2,
	% 	% mark size=0.5pt,
	% ]
	% table[
	% 	x=Rank,
	% 	y=Value,
	% ]
	% \tableAlexaRankBucketsSortedSeRandomHttpWww;

	% \addlegendentry{se.2014-07-10.random.100000-http-www}

	\addplot[
		% only marks,
		orange,
		% opacity=0.2,
		% mark size=0.5pt,
	]
	table[
		x=Rank,
		y=Value,
	]
	\tableAlexaRankBucketsSortedAlexaRandomHttpWww;

	\addlegendentry{alexa.2014-09-01.random.10000-http-www}

	\addplot[
		% only marks,
		green,
		% opacity=0.2,
		% mark size=0.5pt,
	]
	table[
		x=Rank,
		y=Value,
	]
	\tableAlexaRankBucketsSortedAlexaTopHttpWww;

	\addlegendentry{alexa.2014-09-01.top.10000-http-www}

	\addplot[
		% only marks,
		red,
		% opacity=0.2,
		% mark size=0.5pt,
	]
	table[
		x=Rank,
		y=Value,
	]
	\tableAlexaRankBucketsSortedAlexaTopSeHttpWww;

	\addlegendentry{alexa.2014-09-01.top.se.10000-http-www}

	\addplot[
		% only marks,
		purple,
		% opacity=0.2,
		% mark size=0.5pt,
	]
	table[
		x=Rank,
		y=Value,
	]
	\tableAlexaRankBucketsSortedAlexaTopDkHttpWww;

	\addlegendentry{alexa.2014-09-01.top.dk.10000-http-www}
\end{semilogxaxis}
\end{tikzpicture}

\end{document}